### 2017-04-17 paper/oopsla17-submission/intro2.tex
small tweaks thanks to abaskt, gleissen, and atondawal

\mypara{A Data-Driven Approach}
%
In this paper, we introduce \toolname
\footnote{``Numeric Analysis of Type Errors''; any resemblance to persons living or dead is purely coincidental.}
a \emph{data-driven} approach to error
localization based on supervised
learning (see \citealt{Kotsiantis2007-pj} for a survey).
%
\toolname analyzes a large corpus
of training data --- pairs of ill-typed
programs and their subsequent fixes ---
to automatically \emph{learn a model}
of where errors are most likely to
be found.
%
Given a new ill-typed program,
\toolname simply executes the model
to generate a list of potential
blame assignments ranked by likelihood.
%
We evaluate \toolname by comparing its
precision against the state-of-the-art
on a set of over 4,500 ill-typed \ocaml
programs drawn from introductory
programming classes.
%
We show that, when restricted to a
\emph{single} prediction, \toolname's data-driven
model is able to correctly predict
the exact sub-expression that should
be changed \HiddenFhTopOne\% of the time,
\ToolnameWinOcaml points higher than \ocaml and
\ToolnameWinSherrloc points higher than the state-of-the-art
\sherrloc tool.
%
Furthermore, \toolname's accuracy surpasses
\HiddenFhTopTwo\% when we consider the top \emph{two}
locations and reaches \HiddenFhTopThree\% if consider
the top \emph{three}.
%
We achieve these advances by identifying
and then solving three key challenges.

@@

Our \emph{second contribution} (\autoref{sec:learning})
solves this problem with a simple, yet expressive, representation called
a \emph{Bag-of-Abstracted-Terms (BOAT)} wherein
each program is represented by the \emph{bag}
or multiset of (sub-)~terms that appears inside
it; and further, each (sub-)~term is \emph{abstracted}
as a feature vector comprising the numeric values
returned by \emph{feature abstraction} functions
that applied to the term.
%
We can model the blame labels with a simple feature
abstraction (\eg is-changed-in-fix).
%
We can even account for \emph{contextual} information
like the features of the parent- and child- terms by
\emph{concatenating} the feature vectors of each term
with those of its parent and children.
%
We have found this representation to be particularly
convenient as it gives us flexibility in modeling the
syntactic and semantic structure of programs while
retaining compatibility with off-the-shelf classifiers,
in contrast to, \eg, \citet{Raychev2015-jg}, who had
to develop their own variants of classifiers to obtain
their results.

\mypara{Challenge 3: Training Precise Classifiers}
%
Finally, the last and most important challenge is to
use the labeled BOAT data to train classifiers that
are capable of \emph{precisely} pinpointing the errors
in real programs.
%
The key here is to find the right set of feature
abstractions and classification algorithms that
lead to precise predictions.
%
Fortunately, our BOAT model allows us a great deal of
latitude in our choice of features.
%
We can use abstraction functions to capture different
aspects of a term ranging from
%
syntactic features (\eg is-a-data constructor, is-a-literal,
is-an-arithmetic-operation, is-a-function-application, \etc),
%
to semantic features captured by the type system (\eg is-a-list,
is-an-integer, is-a-function, \etc).

>>>>

\mypara{A Data-Driven Approach}
%
In this paper, we introduce \toolname
\footnote{``Numeric Analysis of Type Errors''; any resemblance to persons living or dead is purely coincidental.}
a \emph{data-driven} approach to error
localization based on supervised
learning (see \citealt{Kotsiantis2007-pj} for a survey).
%
\toolname analyzes a large corpus
of training data --- pairs of ill-typed
programs and their subsequent fixes ---
to automatically \emph{learn a model}
of where errors are most likely to
be found.
%
Given a new ill-typed program,
\toolname simply executes the model
to generate a list of potential
blame assignments ranked by likelihood.
%
We evaluate \toolname by comparing its
precision against the state-of-the-art
on a set of over 4,500 ill-typed \ocaml
programs drawn from introductory
programming classes.
%
We show that, when restricted to a
\emph{single} prediction, \toolname's data-driven
model is able to correctly predict
the exact sub-expression that should
be changed \HiddenFhTopOne\% of the time,
\ToolnameWinOcaml points higher than \ocaml and
\ToolnameWinSherrloc points higher than the state-of-the-art
\sherrloc tool.
%
Furthermore, \toolname's accuracy surpasses
\HiddenFhTopTwo\% when we consider the top \emph{two}
locations and reaches \HiddenFhTopThree\% if we consider
the top \emph{three}.
%
We achieve these advances by identifying
and then solving three key challenges.

@@

Our \emph{second contribution} (\autoref{sec:learning})
solves this problem with a simple, yet expressive, representation called
a \emph{Bag-of-Abstracted-Terms (BOAT)} wherein
each program is represented by the \emph{bag}
or multiset of (sub-)~terms that appears inside
it; and further, each (sub-)~term is \emph{abstracted}
as a feature vector comprising the numeric values
returned by \emph{feature abstraction} functions
applied to the term.
%
% We can model the blame labels with a simple feature
% abstraction (\eg is-changed-in-fix).
%
We can even recover \emph{contextual} information
from the parent and child terms by
\emph{concatenating} the feature vectors of each term
with those of its parent and children
(within a fixed window).
%
We have found this representation to be particularly
convenient as it gives us flexibility in modeling the
syntactic and semantic structure of programs while
retaining compatibility with off-the-shelf classifiers,
in contrast to, \eg, \citet{Raychev2015-jg}, who had
to develop their own variants of classifiers to obtain
their results.

\mypara{Challenge 3: Training Precise Classifiers}
%
Finally, the last and most important challenge is to
use our BOAT representation to train classifiers that
are capable of \emph{precisely} pinpointing the errors
in real programs.
%
The key here is to find the right set of feature
abstractions to model type errors,
and classification algorithms that
lead to precise blame assignments.
%
Fortunately, our BOAT model allows us a great deal of
latitude in our choice of features.
%
We can use abstraction functions to capture different
aspects of a term ranging from
%
syntactic features (\eg is-a-data-constructor, is-a-literal,
is-an-arithmetic-operation, is-a-function-application, \etc),
%
to semantic features captured by the type system (\eg is-a-list,
is-an-integer, is-a-function, \etc).
%
We can similarly model the blame labels with a simple feature
abstraction (\eg is-changed-in-fix).

### 2017-04-17 paper/oopsla17-submission/overview.tex
small tweaks thanks to abaskt, gleissen, and atondawal

\mypara{Solution: Bags-of-Abstracted-Terms}
%
We present a new representation of programs
that draws inspiration from the theory of
abstract interpretation \citep{CousotCousot77}.
%
Our representation as parameterized by a
set of \emph{feature abstraction} functions,
(abbreviated to feature abstractions)
$f_1, \ldots, f_n$, that map terms to a
numeric value (or just $\{0, 1\}$ to
encode a boolean property).
%
Given a set of feature abstractions, we
can represent a single program's AST as
a \emph{bag-of-abstracted-terms} (BOAT)
by:
%
(1)~decomposing the AST (term) $t$ into
    a \emph{bag} of its constituent sub-trees
    (terms) $\{t_1,\ldots,t_m\}$; and then
%
(2)~representing each sub-term $t_i$
    with the $n$-dimensional
    vector $[f_1(t_i),\ldots, f_n(t_i)]$.
%
Our choice of working with ASTs is natural
as that is the representation over which the
type-checker operates.

@@

\mypara{Syntax and Type Features}
%
We start by observing that
at the very least, the
classifier should be able
to distinguish between the
|[]| and |+| expressions
in \autoref{fig:sumList}
because they represent
different \emph{syntactic}
expression forms.
%
We model this by
introducing feature
abstractions of the form
is-nil, is-plus, \etc for
each of a fixed number of
primitive literal terms.
%
Modeling the syntactic class of an
expression gives the classifier a
basic notion of the relative
frequency of blame assignment
for the various program elements,
\ie perhaps |[]| is
\emph{empirically} more
likely to be blamed than |+|.
%
Similarly, we can model
the \emph{type} of each
sub-expression with features
of the form is-int, is-bool, \etc.
%
We will discuss handling
arbitrary, user-defined types
in~\autoref{sec:discussion}.

>>>>

\mypara{Solution: Bags-of-Abstracted-Terms}
%
We present a new representation of programs
that draws inspiration from the theory of
abstract interpretation \citep{CousotCousot77}.
%
Our representation is parameterized by a
set of \emph{feature abstraction} functions,
(abbreviated to feature abstractions)
$f_1, \ldots, f_n$, that map terms to a
numeric value (or just $\{0, 1\}$ to
encode a boolean property).
%
Given a set of feature abstractions, we
can represent a single program's AST as
a \emph{bag-of-abstracted-terms} (BOAT)
by:
%
(1)~decomposing the AST (term) $t$ into
    a \emph{bag} of its constituent sub-trees
    (terms) $\{t_1,\ldots,t_m\}$; and then
%
(2)~representing each sub-term $t_i$
    with the $n$-dimensional
    vector $[f_1(t_i),\ldots, f_n(t_i)]$.
%
Our choice of working with ASTs is natural
as that is the representation over which the
type-checker operates.

@@

\mypara{Syntax and Type Features}
%
We start by observing that
at the very least, the
classifier should be able
to distinguish between the
|[]| and |+| expressions
in \autoref{fig:sumList}
because they represent
different \emph{syntactic}
expression forms.
%
We model this by
introducing feature
abstractions of the form
is-|[]|, is-|+|, \etc, for
each of a fixed number of
expression forms.
%
Modeling the syntactic class of an
expression gives the classifier a
basic notion of the relative
frequency of blame assignment
for the various program elements,
\ie perhaps |[]| is
\emph{empirically} more
likely to be blamed than |+|.
%
Similarly, we can model
the \emph{type} of each
sub-expression with features
of the form is-|int|, is-|bool|, \etc.
%
We will discuss handling
arbitrary, user-defined types
in~\autoref{sec:discussion}.
